---
title: "ESS 575: Probability Lab 2 - Probability Distributions"
author: "Team England" 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: false
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding){ 
    out_dir <- '../';
    rmarkdown::render(inputFile, encoding = encoding, output_file=file.path(dirname(inputFile), out_dir, 'ProbLab_2_England.pdf')) 
  })
---

Team England:

  - Caroline Blommel 
  - Carolyn Coyle
  - Bryn Crosby
  - George Woolsey
  
cblommel@mail.colostate.edu, carolynm@mail.colostate.edu, brcrosby@rams.colostate.edu, george.woolsey@colostate.edu


```{r setup, include=FALSE}
# load packages
library(tidyverse)
library(lubridate)
library(viridis)
library(scales)
library(gridExtra)
# set seed
set.seed(10)
```

# Question 1

We commonly represent the following general framework for linking models to data:

\begin{equation}
\tag{1}
[y_{i} \mid g(\theta , x_{i}), \sigma^{2}]
\end{equation}

which represents the probability of obtaining the observation $y_{i}$ given that our model predicts the mean of a distribution $g(\theta , x_{i})$ with variance $\sigma^{2}$. Assume we have count data. What distribution would be a logical choice to model these data? Write out a model for the data.

```{r, warning=F, message=F, fig.width = 10, fig.height = 10, eval=FALSE, include}
# lambda
lambda <- seq(0.25, 4, 0.25)

# time
t <- seq(1, 30, 1)

# x
x <- numeric(length(t))
x[1] <- 0.01

# set up function
fn_x_t <- function(x, t, lambda) {
  x_t <- lambda * x[t-1] * (1 - x[t-1])
  return(x_t)
}

# set up plot grid
plts <- list()

#loop through lambda values
for (my_l in 1:length(lambda)) {
  # assign values of x by t
  for (my_t in 2:length(t)) {
    x[my_t] <- fn_x_t(x, my_t, lambda[my_l])
  }
  
  # set up temp data
  dta <- data.frame(
    t
    , x
  )
  # plot
  plts[[my_l]] <-
    ggplot(dta, aes(x = t, y = x)) +
      geom_line(
        size = 1
        , alpha = 0.8
      ) +
      labs(
        title = bquote(paste(
            lambda
            , " = "
            , .(lambda[my_l])
          ))
      ) +
      xlab("time") +
      ylab("x") +
      theme_bw() +
      theme(
        legend.position="none"
        , plot.title = element_text(size = 9)
        , axis.text =  element_text(size = 8)
        , axis.title =  element_text(size = 8)
      )
}

# pass plots to grid extra package
do.call(gridExtra::grid.arrange, plts)
```

```{r, warning=F, message=F, fig.width = 8, fig.height = 6, eval=FALSE, include=FALSE}
rm(list=ls())
# lambda
lambda <- seq(1, 4, 0.05)

# time
t <- seq(1, 100, 1)

# x
x <- numeric(length(t))
x[1] <- 0.01

# set up function
fn_x_t <- function(x, t, lambda) {
  x_t <- lambda * x[t-1] * (1 - x[t-1])
  return(x_t)
}

# set up empty vector
  vect_lamda <- c()
  vect_xt <- c()
  vect_t <- c()

#loop through lambda values
for (my_l in 1:length(lambda)) {
  
  # assign values of x by t
  for (my_t in 2:length(t)) {
    x[my_t] <- fn_x_t(x, my_t, lambda[my_l])
  }
  
  # add vectors together
  vect_lamda <- c(vect_lamda, rep(lambda[my_l], each=length(t[51:100])))
  vect_xt <- c(vect_xt, x[51:100])
  vect_t <- c(vect_t, t[51:100])
}
# put vectors together  
fnl_matrix <- matrix(c(vect_lamda, vect_xt, vect_t), ncol = 3)
fnl_data <- data.frame(vect_lamda, vect_xt, vect_t)

# plot
  ggplot(fnl_data, aes(x = vect_lamda, y = vect_xt)) +
    geom_point(
      alpha = 0.8
    ) +
    labs(
      title = "Bifurcation Diagram"
    ) +
    xlab(expression(lambda)) +
    ylab(expression(x["t"])) +
    theme_bw() +
    theme(
      legend.position="none"
    )

```

